---
title: "Spatial_analysis_RASTER_JS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Read in landcover raster 
```{r}
library(raster)

lc_3338 <- raster("/home/shares/scientist/ds-workshop/ak_nlcd_2011_landcover_50m_3338.tif")
```

```{r}
plot(lc_3338)
plot(pop_3338, add=TRUE)
```


#Extracting most frequent land cover
```{r}
raster_points <- extract(lc_3338, pop_3338, buffer = 500, fun=modal)
raster_points
```

```{r}
pop_joined$land_cover <- raster_points
pop_joined
```


# Aggregate LC categories (load legend, aggregate attributes etc.)
```{r}
# Read the land cover legend
Legend_lc <- read.csv("/home/shares/scientist/ds-workshop/legend_ak_nlcd_2011.csv", stringsAsFactors = FALSE)
Legend_lc
```

```{r}
# It is a lot of categories, let us consolidate this
Legend_lc <- Legend_lc %>%
  mutate(main_lc = ifelse(ID %in% 40:49, "Forest",
                              ifelse(ID %in% 20:29, "Urban",
                                     ifelse(ID %in% 50:59, "Shrub",
                                             ifelse(ID %in% 70:79, "Grass",
                                                     ifelse(ID %in% 80:89, "Crops",
                                                             ifelse(ID %in% 90:99, "Wetland", Land.Cover.Type)
                                                             )
                                                     )
                                             )
                                     )
                              )
  )


# Join the LC categories to the population data
pop_3338_cover <- left_join(pop_joined, Legend_lc, by=c("land_cover"="ID")) %>% 
  dplyr::select(-Red, -Green, -Blue, -Land.Cover.Type)

pop_3338_cover
```


#Plot lc
```{r}
# Create color palette by keeping last color of each group
pal <- Legend_lc %>% 
  group_by(main_lc) %>% 
  slice(n()) %>% # Keeping the last color of the groups
  ungroup %>% 
  arrange(ID) %>%
  mutate(color_hex = rgb(Red, Green, Blue, max = 255)) %>% 
  dplyr::select(main_lc, color_hex)

# turn pallete into a list for plotting
pal_list <- pal$color_hex
names(pal_list) <- pal$main_lc


# Plot by region
ggplot(pop_3338_cover, aes(region, population, fill = main_lc)) +
    geom_col() +
    scale_y_continuous(labels = comma) +
    scale_fill_manual(values = pal_list) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Land cover type by region and population") +
    labs(fill = "", y = "Population", x = "")
```

#Map - colour by lc, dot size by pop
```{r}
ggplot() +
    geom_sf(data = ak_regions_3338, aes(), color = "black") +
    geom_sf(data = pop_3338_cover, aes(color = main_lc,
                                       size = population), show.legend = "point") +
    scale_size_continuous(guide = F) +
    scale_color_manual(values = pal_list) +
    theme_bw() +
    theme(legend.position = "bottom", legend.title = element_blank())
```


#Cropping the Copper River (equivalent of EXTRACT BY MASK)
```{r}
copper_region_3338 <- ak_regions_3338 %>%
  filter(region =="Copper River")
```

```{r}
copper_lc_3338 <- crop(lc_3338, copper_region_3338)
plot(copper_lc_3338)
```


#Create a mask for forest land cover type - need to create a function
```{r}
forest_mask <-function(x){
  x[x<40 | x>49] <- 0
  x[x>=40 & x<=49] <- 1
  x
}
```


# Calculate new pixel values (e.g. )
```{r}
copper_forested_3338 <- calc(copper_lc_3338, forest_mask)
plot(copper_forested_3338)
```


#Calculate amount of forest around urban centres in Copper River
```{r}
copper_pop_3338 <- pop_3338_cover %>% 
  filter(region == "Copper River")

plot(copper_pop_3338)
```

```{r}
forested_count <- extract(copper_forested_3338, copper_pop_3338, buffer=500, fun=sum)
```

```{r}
copper_pop_3338$forest_cov <- 100*forested_count / (20*20)
head(copper_pop_3338)
```

```{r}
ggplot() +
    geom_sf(data = copper_region_3338, aes(), color = "black") +
    geom_sf(data = copper_pop_3338, aes(color = forest_cov,
                                       size = population)) +
    scale_size_continuous(guide = F) +
    scale_colour_gradientn(colours = terrain.colors(10, rev=TRUE)) +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs(color = "Percent Forest Cover")
```

